---
title: "Module 3"
author: "Terry Zhou"
date: "3/27/2022"
output: pdf_document
---

```{r setup, include=FALSE, fig.align = 'center', fig.width = 8, fig.height=5}
knitr::opts_chunk$set(echo = TRUE)
dat <- read.csv("module3.csv", header = FALSE)
```

# 1. Warm Up

### 1. Scatterplot

```{r}
library(scatterplot3d)
scatterplot3d(dat$V1, dat$V2, dat$V3, 
              xlab = "V1", ylab = "V2", zlab = "V3",
              main = "3D Scatterplot", color = "blue")
```

### 2. Histogram

```{r}
hist(dat$V3, main = "Histogram of V3", xlab = "V3", col = "hotpink")
```

### 3. Line graph

```{r}
plot(dat$V1, type = "l", col = "orange", ylim = c(0, 0.25), lwd = 2, 
     xlab = "Index", ylab = "Value", main = "Values of Each Variable")
lines(dat$V2, type = "l", col = "blue", lwd = 2)
lines(dat$V3, type = "l", col = "red", lwd = 2)
legend(1, 0.09, legend = c("V1", "V2", "V3"),
       col = c("orange", "blue", "red"), lty = 1, lwd = 2,
       title = "Legend")
```

\newpage

# 2. Data Visualization and Analysis on a Dataset 1

### 2.

```{r}
data(iris)
str(iris)
apply(is.na(iris), 2, which)
summary(iris)
```
The dataset contains 150 observations of 5 variables. The `Sepal.Length`, `Sepal.Width`, `Petal.Length`, and `Petal.Width` variables are all numerical. The `Species` variable is a factor with three levels. There are no missing values. 

### 3.

```{r}
summary(subset(iris, Species == "setosa"))
summary(subset(iris, Species == "versicolor"))
summary(subset(iris, Species == "virginica"))
```

### 4. 
```{r}
library(ggplot2)
library(gridExtra)
plotlist <- list()
colors <- c("forestgreen", "dodgerblue", "firebrick3", "goldenrod1")
for (i in colnames(iris)[-5]){
  a <- (strsplit(i, split = "[.]"))[[1]][1]
  b <- (strsplit(i, split = "[.]"))[[1]][2]
  index <- which(colnames(iris) == i)
  plot <- ggplot(data = iris, aes_string(y = i)) + 
    geom_boxplot(fill = colors[index]) +
    theme_bw() + ylab(paste(a, b, "(cm)", sep = " ")) + 
    ggtitle(paste("Boxplot of", a, b, sep = " ")) + 
    theme(axis.text.x = element_blank(), axis.ticks = element_blank())
    plotlist[[i]] <- plot
}
grobs <- arrangeGrob(grobs = plotlist)
grid.arrange(grobs)
```

Only the sepal width plot shows outliers. 

### 5. 

```{r}
histlist <- list()
colors <- c("forestgreen", "dodgerblue", "firebrick3", "goldenrod1")
for (i in colnames(iris)[-5]){
  a <- (strsplit(i, split = "[.]"))[[1]][1]
  b <- (strsplit(i, split = "[.]"))[[1]][2]
  index <- which(colnames(iris) == i)
  plot <- ggplot(data = iris, aes_string(x = i)) + 
    geom_histogram(fill = colors[index]) +
    theme_bw() + xlab(paste(a, b, "(cm)", sep = " ")) + ylab("Frequency") + 
    ggtitle(paste("Histogram of", a, b, sep = " "))
    histlist[[i]] <- plot
}
grobs.hist <- arrangeGrob(grobs = histlist)
grid.arrange(grobs.hist)
```

None of the distributions are normal, however the histogram of sepal width seems to be the most normal. There are gaps in the histogram for all four variables. The sepal width, petal length, and petal width histograms are unimodal, however the sepal length histogram seems multimodal. The sepal length, petal length, and petal width histograms all seem to be skew right. 

### 6.

```{r}
library(ggplot2)
library(GGally)
ggpairs(iris[, 1:4], aes(colour = iris$Species, alpha = 0.4)) + theme_bw()
```

There is a distinction between the three species for the petal width and petal length variables. 

### 7.

```{r, warning=FALSE}
cor.matrix <- cor(iris[, c(1:4)])
cor.matrix
library(gplots)
heatmap.2(cor.matrix, cexRow = 1, cexCol = 1,
          Rowv = FALSE, Colv = FALSE, trace = c("none"), density.info = "none",
          lhei = c(1, 3), lwid = c(0.5, 1.5), margins = c(7, 7))
```

There are strong correlation between Sepal Length and Petal Length, Sepal Length and Petal Width, and Petal Length and Petal Width. This is because the correlation coefficients for each of these pairs is greater than 0.8. 

### 8. 

```{r}
sepal.length.a <- t.test(iris$Sepal.Length, conf.level = 0.95)
sepal.length.a
```

We are 95% confident that the mean sepal length of all observations is between 5.724154 cm and 5.962512 cm. The p-value of the one sample t-test is is less than 0.05, thus we can reject the null hypothesis and confirm that the mean sepal length is not equal to 0. 

```{r}
sepal.width.b <- t.test(iris$Sepal.Width, conf.level = 0.95, 
                         alternative = "greater", mu = 4)
sepal.width.b
```

The null hypothesis is that the mean sepal width is equal to 4 cm. The alternative hypothesis is that the mean sepal width is greater than 4 cm. The one sample t-test gives a p-value of 1, which is greater than our alpha level of 0.05. Thus, we fail to reject the null hypothesis; there is not enough evidence to show that the mean sepal width is greater than 4 cm. The sample estimate for the mean sepal width is 3.057333. We are 95% confident that the mean sepal width is greater than 2.998429.

```{r}
iris1 <- subset(iris, Species == "setosa" | Species == "versicolor")
var.test(Petal.Length ~ Species, iris1, alternative = "two.sided")
```

The p-value for the F-test to compare the variances, which is less than our alpha value of 0.05. Thus, we can fail to reject the null hypothesis and conclude that there is a difference in the variance of the petal length in iris setosa and iris versicolor. Thus, we cannot assume that they have equal variances

```{r}
t.test(Petal.Length ~ Species, iris1, alternative = "two.sided",
         conf.level = 0.99, var.diff = TRUE)
```

The two sample t-test gives a p-value of less than 2.2e-16, which is less than the alpha level of 0.01. Thus, we can reject the null hypothesis and conclude that there is a difference in the mean petal length between the iris setosa and iris versicolor species. 

```{r}
petal.length.aov <- aov(Petal.Length ~ Species, data = iris)
summary(petal.length.aov)
```

The ANOVA gives a p-value of less than 2e-16 for the species variable. Thus, we can reject the null hypothesis and conclude, at a alpha = 0.05 significance level, that the average petal length is different between the three species categories. 

### 9.

```{r}
setosa <- subset(iris, Species == "setosa")
versicolor <- subset(iris, Species == "versicolor")
virginica <- subset(iris, Species == "virginica")
which(setosa$Sepal.Width == max(setosa$Sepal.Width)) 
max(setosa$Sepal.Width)
which(versicolor$Sepal.Width == max(versicolor$Sepal.Width)) + 50
max(versicolor$Sepal.Width)
which(virginica$Sepal.Width == max(virginica$Sepal.Width)) + 100
max(virginica$Sepal.Width)
```

Observation 16 has the highest septal width value (4.4 cm) for iris setosa, 86 has the highest value (3.4 cm) for iris versicolor, and 118 and 132 have the highest values (3.8 cm) for iris virginica.

### 10. 

```{r}
median(setosa$Sepal.Length)
median(versicolor$Sepal.Length)
median(virginica$Sepal.Length)
```

The median sepal lengths are 5 cm, 5.9 cm, and 6.5 cm, for iris setosa, iris versicolor, and iris virginica, respectivley. 

### 11. 

```{r}
lm <- lm(iris$Petal.Width ~ iris$Petal.Length)
plot(x = iris$Petal.Length, y = iris$Petal.Width, 
     xlab = "Petal Length (cm)", ylab = "Petal Width (cm)",
     main = "Petal Width As A Function of Petal Length", col = "deeppink3",
     pch = 18)
abline(lm, col = "olivedrab4", lwd = 2)
predict.2.5 <- as.numeric(lm$coefficients[1] + (lm$coefficients[2] * 2.5))
predict.2.5
```

There is a strong positive relationship between petal width and petal length; as petal length increases, so does petal width. At a petal length of 2.5 cm, we can expect the petal width to be 0.676313 cm. 

\newpage

# 2. Data Visualization and Analysis on a Dataset 2

### 1. 

```{r}
heart <- read.csv("processed.cleveland.data", header = FALSE)
colnames(heart) <- c("age", "sex", "cp", "trestbps", "chol", "fbs", "restecg", 
                     "thalach", "exang", "oldpeak", "slope", "ca", 
                     "thal", "num")
str(heart)
apply(is.na(heart), 2, which)
summary(heart)
```

There are 14 variables. All of the variables except `ca` and `thal` are numeric; `ca` and `thal` are both characters. 

### 3. 

```{r}
hist(heart$trestbps, main = "Histogram of Resting Blood Pressure",
     xlab = "Resting Blood Pressure (mm Hg)", col = "deeppink3")
hist(log(heart$trestbps), 
     main = "Histogram of Log Transformed Resting Blood Pressure",
     xlab = "Resting Blood Pressure (log(mm Hg))", col = "deeppink3")
```

The resting blood pressure is slightly right skew. To reduce the skew, I would take the log transformation of the variable. 

### 4.

```{r}
library(dplyr)
heart <- heart %>%
  mutate(age_group = case_when((age >= 20 & age <= 30) ~ "20-30",
                               (age >= 31 & age <= 40) ~ "31-40",
                               (age >= 41 & age <= 50) ~ "41-50",
                               (age >= 51 & age <= 60) ~ "51-60",
                               (age >= 61 & age <= 70) ~ "61-70",
                               (age >= 71 & age <= 80) ~ "71-80"))
age.tot <- heart %>%
  group_by(age_group) %>%
  summarize(tot = n())
cp.tot <- heart %>%
  group_by(cp) %>%
  summarize(tot = n())
tot <- heart %>%
  group_by(age_group, cp) %>%
  summarize(tot = n())
mat <- matrix(nrow = 7, ncol = 5)
mat[1, 4] <- 0 
mat[1, 2] <- 0 
mat[1, 1] <- 0
mat[1, 3] <- tot$tot[1]
mat[2, 4:1] <- tot$tot[2:5]
mat[3, 4:1] <- tot$tot[6:9]
mat[4, 4:1] <- tot$tot[10:13]
mat[5, 4:1] <- tot$tot[14:17]
mat[6, 4] <- 0
mat[6, 3:1] <- tot$tot[18:20]
mat[1:6, 5] <- age.tot$tot
mat[7, 4:1] <- cp.tot$tot
mat[7, 5] <- nrow(heart)
rownames(mat) <- c(age.tot$age_group, "Total")
colnames(mat) <- c("Asymptomatic", "Non-Anginal Pain", "Atypical Angina", 
                   "Typical Angina", "Total")
mat
```


